---
title: "Exercises 2"
author: "Wooyoung Jang"
date: "6/16/2021"
output: html_document
---

## Exercise 2.1
Note that $u \leq F(F^-(u))$ and $x \geq F^-(F(x))$. Show that $P(F^-(U) \leq x) \leq P(U \leq F(x)) \leq P(F^-(U) \leq x)$. The result directly follows.

## Exercise 2.2
### (a)
```{r}
supp = seq(-12, 12, 0.01)
N = 10000
U = runif(N)
f = dlogis
Finv = function(x) { -log((1-x)/x) }
X = Finv(U); X = X[-12 < X & X < 12]
Y = rlogis(N); Y = Y[-12 < Y & Y < 12]

par(mfrow=c(1,2))
hist(X, freq=FALSE, breaks=seq(-12, 12, 0.5), main="Logistic from Uniform")
lines(supp, f(supp), col="red", lwd=2)
hist(Y, freq=FALSE, breaks=seq(-12, 12, 0.5), main="Logistic from R")
lines(supp, f(supp), col="red", lwd=2)
```

### (b)
```{r}
supp = seq(-12, 12, 0.01)
N = 10000
U = runif(N)
f = dcauchy
Finv = function(x) { tan(pi * (x - 1/2)) }
X = Finv(U); X = X[-12 < X & X < 12]
Y = rcauchy(N); Y = Y[-12 < Y & Y < 12]

par(mfrow=c(1,2))
hist(X, freq=FALSE, breaks=seq(-12, 12, 0.5), main="Cauchy from Uniform")
lines(supp, f(supp), col="red", lwd=2)
hist(Y, freq=FALSE, breaks=seq(-12, 12, 0.5), main="Cauchy from R")
lines(supp, f(supp), col="red", lwd=2)
```

## Exercise 2.3
### (a)
Note that $\mathbb{E}[U] =\frac{1}{2}$ and $Var(U) = \frac{1}{12}$ if $U \sim U(0, 1)$.

### (b)
```{r}
N = 10000
test1 = function(N) {
  U = matrix(runif(12 * N), nrow=12) - 0.5
  Z = apply(U, 2, sum)
}
test2 = function(N) {
  U1 = runif(N)
  U2 = runif(N)
  X1 = sqrt(-2 * log(U1)) * cos(2 * pi * U2)
}
par(mfrow=c(1, 2))
hist(test1(N), freq=FALSE, breaks=seq(-6, 6, 0.2), main="CLT-normal generator")
hist(test2(N), freq=FALSE, breaks=seq(-6, 6, 0.2), main="Box-Muller algorithm")

system.time(test1(N)); system.time(test2(N))
```

### (c)
```{r}
test3 = function(N) {
  rnorm(N)
}
hist(test3(N), freq=FALSE, breaks=seq(-6, 6, 0.2), main="Normal from R")
system.time(test3(N))
```

## Exercise 2.4
### (a), (b)
Omit.

### (c)
```{r}
N = 10000
Sigma = cov(matrix(rnorm(30), nrow=10))

test1 = function(N) {
  A = t(chol(Sigma))
  X = matrix(runif(3 * N), nrow=N)
  X = X %*% A
}
test2 = function(N) {
  library(mnormt)
  rmnorm(N, varcov=Sigma)
}

system.time(test1(N)); system.time(test2(N))
```


## Exercise 2.5
- $P(U \leq f(Y)/Mg(Y)) = \int_{-\infty}^\infty \int_0^{f(y)/Mg(y)} du \cdot g(y) dy = \int_{-\infty}^\infty \frac{f(y)}{Mg(y)} \cdot g(y) dy = \frac{1}{M}.$
- Let $\tilde{f} / \tilde{g} = cf/g$ for some $c$. Then we can observe that
$\mathbb{E}[1_{(U < \tilde{f}(Y) / \tilde{M} \tilde{g}(Y))}] = \int_{-\infty}^\infty \int_0^{\tilde{f}(y)/\tilde{M}\tilde{g}(y)} du \cdot g(y)dy = \int_{-\infty}^\infty cf(y)/\tilde{M}g(y) \cdot g(y) dy = c/\tilde{M}$.
Note that we can estimate $\mathbb{E}[1_{(U < \tilde{f}/\tilde{M}\tilde{g})}]$ by the empirical acceptance rate. Since we know $\tilde{M}$, we can estimate $c$ as well.



## Exercise 2.12
### (a)
```{r}
a = 3
b = 5
N = 10^4

test1 = function(N, a, b) {
  U = matrix(runif(a * N), nrow=a)
  X = b * apply(-log(U), 2, sum)
}
test2 = function(N, a, b) {
  X = rgamma(N, a, scale=b)
}
system.time(test1(N, a, b)); system.time(test2(N, a, b))

test1 = function(N, a, b) {
  U = matrix(runif((a+b) * N), nrow=a+b)
  X = -log(U)
  X = apply(X[1:a,], 2, sum) / apply(X, 2, sum)
}
test2 = function(N, a, b) {
  rbeta(N, a, b)
}
system.time(test1(N, a, b)); system.time(test2(N, a, b))
```

### (b), (c)
Omit.








## Exercise 2.14

### (a)
Omit

### (b)
```{r}
N = 1000
r = 10

par(mfrow=c(1,3))
for (p in c(0.01, 0.1, 0.5)) {
  tmean = r*(1-p)/p
  tstd = sqrt(r*(1-p)/p^2)
  mrange = round(seq(max(0, tmean - 3*tstd), tmean + 3*tstd, 1))
  prob = pnbinom(mrange, r, p)
  X = rep(0, N)
  
  for (i in 1:N) {
    u = runif(1)
    X[i] = mrange[1] + sum(prob < u)
  }
  
  hist(X, freq=FALSE, main=paste("Negative Binomial with p=", p, sep=""))
  lines(dnbinom(seq(1, max(X)), r, p), col="sienna", lwd=2)
}
```

### (c)
```{r}
N = 1000

par(mfrow=c(1, 3))
for (p in c(0.001, 0.01, 0.5)) {
  tmean = -(1-p)/(p*log(p))
  tvar = -(1-p)/(p^2*log(p)) - tmean^2
  tstd = sqrt(tvar)
  mrange = round(seq(max(1, tmean - 3*tstd), tmean + 3*tstd, 1))
  density = function(x, p) {
    -(1-p)^x/(x*log(p))
  }
  prob = cumsum(density(mrange, p))
  X = rep(1, N)
  
  for (i in 1:N) {
    u = runif(1)
    X[i] = mrange[1] + sum(prob < u)
  }
  
  M = max(X)
  
  hist(X, freq=FALSE, main=paste("Logarithmic Series with p=", p, sep=""))
  lines(density(seq(1, M), p), col="sienna", lwd=2)
}
```

The histogram when $p=0.5$ seems to be an error of R. The first two histogram show that the tail distribution is so heavy, in which it is not unlikable to be happened even outside of three times standard deviation further from the mean.


## Exercise 2.18
```{r}
N = 2500
supp = seq(-3, 3, 0.001)

h = function(x) { exp(-x^2/2)*(sin(6*x)^2 + 3*cos(x)^2*sin(4*x)^2 + 1) }
g = dnorm
M = 5 * sqrt(2 * pi)
X = rnorm(N); X = X[-3 < X & X < 3]
U = runif(length(X))
Y = M * U * g(X)

AcceptX = X[Y <= h(X)]
RejectX = X[Y > h(X)]
rate = length(AcceptX) / N
c = rate * M
AcceptY = Y[Y <= h(X)] / c
RejectY = Y[Y > h(X)] / c
f = function(x) { h(x) / c }

plot(RejectX, RejectY, xlab="", ylab="", pch=20, cex=0.5, col="gray")
points(AcceptX, AcceptY, pch=20, cex=0.5)
lines(supp, g(supp) / rate, col="orange", lwd=2)
lines(supp, f(supp), col="red", lwd=2)
legend("topright", legend=c("g(x)", "f(x)"), col=c("orange", "red"), lty=c(1, 1))
```

## Exercise 2.19
```{r}
N = 2500
supp = seq(-5, 5, 0.001)

f = dnorm
g = function(x) { exp(-abs(x))/2 }
M = sqrt(2 * exp(1) / pi)
X = rexp(N) * (2*rbinom(N, 1, 0.5) - 1); X = X[-5 < X & X < 5]
U = runif(length(X))
Y = M * U * g(X)

AcceptX = X[Y <= f(X)]
RejectX = X[Y > f(X)]
AcceptY = Y[Y <= f(X)]
RejectY = Y[Y > f(X)]

plot(RejectX, RejectY, xlab="", ylab="", pch=20, cex=0.5, col="gray")
points(AcceptX, AcceptY, pch=20, cex=0.5)
lines(supp, g(supp) * M, col="orange", lwd=2)
lines(supp, f(supp), col="red", lwd=2)
legend("topright", legend=c("g(x)", "f(x)"), col=c("orange", "red"), lty=c(1, 1))
```

## Exercise 2.20
### (a)
```{r}
N = 2500
supp = seq(-3, 3, 0.001)

f = dnorm
g = dcauchy
M = sqrt(2*pi / exp(1))
X = rcauchy(N); X = X[-3 < X & X < 3]
U = runif(length(X))
Y = M * U * g(X)

AcceptX = X[Y <= f(X)]
RejectX = X[Y > f(X)]
AcceptY = Y[Y <= f(X)]
RejectY = Y[Y > f(X)]

plot(RejectX, RejectY, xlab="", ylab="", pch=20, cex=0.5, col="gray")
points(AcceptX, AcceptY, pch=20, cex=0.5)
lines(supp, g(supp) * M, col="orange", lwd=2)
lines(supp, f(supp), col="red", lwd=2)
legend("topright", legend=c("g(x)", "f(x)"), col=c("orange", "red"), lty=c(1, 1))
```

### (b)
```{r}
N = 2500
supp = seq(0, 3, 0.001)

f = function(x) { dgamma(x, 4.3, 6.2) }
g = function(x) { dgamma(x, 4, 6) }
f_g = function(x) { f(x)/g(x) }
M = optimize(f_g, c(0, 100), maximum=TRUE)$objective
X = rgamma(N, 4, 6)
U = runif(N)
Y = M * U * g(X)

AcceptX = X[Y <= f(X)]
RejectX = X[Y > f(X)]
AcceptY = Y[Y <= f(X)]
RejectY = Y[Y > f(X)]

plot(RejectX, RejectY, xlim=c(0, 3), xlab="", ylab="", pch=20, cex=0.5, col="gray")
points(AcceptX, AcceptY, pch=20, cex=0.5)
lines(supp, g(supp) * M, col="orange", lwd=2)
lines(supp, f(supp), col="red", lwd=2)
legend("topright", legend=c("g(x)", "f(x)"), col=c("orange", "red"), lty=c(1, 1))
```